---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rollNonlinear

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rollNonlinear)](https://CRAN.R-project.org/package=rollNonlinear)
[![R-CMD-check](https://github.com/rollnonlinear/rollNonlinear/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rollnonlinear/rollNonlinear/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/rollnonlinear/rollNonlinear/branch/main/graph/badge.svg)](https://app.codecov.io/gh/rollnonlinear/rollNonlinear?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`rollNonlinear` computes rolling-window nonlinear time-series metrics —
Hurst exponent, DFA, sample/approximate entropy, Lyapunov exponent, and
correlation dimension — on OHLC data for one or more symbols. It provides an
extensible metric registry, built-in series transforms, and optional parallel
execution via [`mirai`](https://mirai.r-lib.org/).

## Installation

```{r, eval = FALSE}
# CRAN (once released):
install.packages("rollNonlinear")

# Development version from GitHub:
# install.packages("remotes")
remotes::install_github("rollnonlinear/rollNonlinear")
```

## Quick start

```{r example}
library(rollNonlinear)
library(data.table)

set.seed(1L)
n  <- 300L
dt <- data.table(
  symbol = rep(c("AAPL", "MSFT"), each = n),
  time   = rep(
    seq(as.POSIXct("2024-01-02 09:30:00", tz = "UTC"), by = 60, length.out = n),
    2L
  ),
  close  = c(cumsum(rnorm(n)) + 150, cumsum(rnorm(n)) + 370)
)

out <- roll_nonlinear(
  dt,
  metrics     = c("hurst", "dfa"),
  window_size = 128L,
  stride      = 32L
)

head(out)
```

## Parallel execution

```{r parallel, eval = FALSE}
out_par <- roll_nonlinear(
  dt,
  metrics     = c("hurst", "dfa"),
  window_size = 128L,
  stride      = 32L,
  parallel    = TRUE,   # uses mirai daemons
  n_workers   = 2L      # optional: override core count
)
```

## Learn more

- **Vignette**: `vignette("rolling-nonlinear", package = "rollNonlinear")`
- **Reference**: `?roll_nonlinear`, `?register_metric`, `?rn_daemons`
- **pkgdown site**: <https://rollnonlinear.github.io/rollNonlinear/>
