---
title: "Rolling Nonlinear Time-Series Metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rolling Nonlinear Time-Series Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r setup}
library(rollNonlinear)
library(data.table)
```

## 1. Introduction

`rollNonlinear` computes rolling-window nonlinear time-series metrics on
intraday OHLC data. It is designed for quantitative research workflows where
you need to characterise the *dynamical regime* of a price series — e.g.,
trend persistence (Hurst), scaling behaviour (DFA), complexity (sample/
approximate entropy), or chaotic divergence (Lyapunov, correlation dimension)
— over time rather than at a single point.

Key features:

- **Six built-in metrics**: `hurst`, `dfa`, `sample_entropy`, `approx_entropy`,
  `lyapunov`, `corr_dim`.
- **Extensible registry**: add custom metrics via `register_metric()`.
- **Built-in series transforms**: `series_close()`, `series_log_returns()`,
  `series_typical()`, `series_range()`, or any custom `function(dt)`.
- **Multi-symbol dispatch**: accepts a data.table with a `symbol` column and
  processes each symbol independently.
- **Optional parallel execution** via `mirai` for multi-symbol workloads.

## 2. Input format

`roll_nonlinear()` expects a `data.table` (or `data.frame`, which is coerced)
with at minimum:

| Column | Type | Notes |
|--------|------|-------|
| `symbol` | character | Symbol identifier |
| `time` | POSIXct / Date / integer | Strictly increasing *within* each symbol |
| price column(s) | numeric | Determined by the `series` argument |

```{r input-format}
set.seed(1L)
n <- 300L

dt <- data.table(
  symbol = rep(c("AAPL", "MSFT"), each = n),
  time   = c(
    seq(as.POSIXct("2024-01-02 09:30:00", tz = "UTC"), by = 60, length.out = n),
    seq(as.POSIXct("2024-01-02 09:30:00", tz = "UTC"), by = 60, length.out = n)
  ),
  open   = c(rep(150, n), rep(370, n)),
  high   = c(cumsum(rnorm(n, 0, 0.3)) + 151, cumsum(rnorm(n, 0, 0.3)) + 371),
  low    = c(cumsum(rnorm(n, 0, 0.3)) + 149, cumsum(rnorm(n, 0, 0.3)) + 369),
  close  = c(cumsum(rnorm(n)), cumsum(rnorm(n))) + c(rep(150, n), rep(370, n))
)

head(dt)
```

## 3. Basic usage — single symbol, Hurst + DFA

```{r basic, eval = TRUE}
single <- dt[symbol == "AAPL"]

out <- roll_nonlinear(
  single,
  metrics     = c("hurst", "dfa"),
  window_size = 128L,
  stride      = 16L
)

head(out)
```

Each row corresponds to one rolling window. `window_start_idx` and
`window_end_idx` are 1-based row indices into the input symbol's data.

## 4. Multi-symbol dispatch

Pass the full multi-symbol table:

```{r multi}
out_multi <- roll_nonlinear(
  dt,
  metrics     = c("hurst", "dfa"),
  window_size = 128L,
  stride      = 32L
)

out_multi[, .(symbol, time, hurst, dfa_alpha)][1:6]
```

The result is keyed by `(symbol, time)`.

## 5. Custom series transforms

### Built-in: log-returns

```{r log-returns}
lr_out <- roll_nonlinear(
  dt[symbol == "AAPL"],
  series      = series_log_returns(),
  metrics     = "hurst",
  window_size = 128L,
  stride      = 32L
)
head(lr_out[, .(time, hurst)])
```

### Typical price

```{r typical}
tp_out <- roll_nonlinear(
  dt[symbol == "AAPL"],
  series      = series_typical(),
  metrics     = "hurst",
  window_size = 128L,
  stride      = 32L
)
head(tp_out[, .(time, hurst)])
```

### User-defined transform

Any function `function(dt) -> numeric` is accepted:

```{r user-fn}
log_close <- function(dt_chunk) log(dt_chunk[["close"]])

out_logclose <- roll_nonlinear(
  dt[symbol == "AAPL"],
  series      = log_close,
  metrics     = "hurst",
  window_size = 128L,
  stride      = 32L
)
head(out_logclose[, .(time, hurst)])
```

## 6. Metric arguments

Use `metric_args` to override per-metric defaults:

```{r metric-args}
ent_out <- roll_nonlinear(
  dt[symbol == "AAPL"],
  metrics     = c("sample_entropy", "approx_entropy"),
  window_size = 100L,
  stride      = 20L,
  metric_args = list(
    sample_entropy = list(dimension = 2L, r_frac = 0.15),
    approx_entropy = list(dimension = 2L, r_frac = 0.20)
  )
)
head(ent_out[, .(time, sample_entropy, approx_entropy)])
```

> **Note:** `r_frac` scales the tolerance `r` as `r_frac * sd(x)` per
> window (volatility-normalised). For volatility-sensitive analysis,
> pass `r` explicitly instead.

## 7. Parallel execution

For workloads with many symbols, `parallel = TRUE` dispatches each symbol to
a local `mirai` daemon:

```{r parallel, eval = FALSE}
# One-time setup (optional — rollNonlinear manages daemons automatically)
rn_daemons(4L)

out_par <- roll_nonlinear(
  dt,
  metrics     = c("hurst", "dfa"),
  window_size = 128L,
  stride      = 32L,
  parallel    = TRUE
)

rn_stop_daemons()
```

### Daemon management

| Scenario | Behaviour |
|----------|-----------|
| No daemons running | Package starts daemons (using `parallelly::availableCores(omit=1)` by default), then tears them down on exit. |
| Daemons already running (user-managed) | Package uses existing daemons and does **not** stop them. |

Use `n_workers` to override the automatic core count:

```{r parallel-workers, eval = FALSE}
roll_nonlinear(dt, parallel = TRUE, n_workers = 2L)
```

### When parallel pays off

Parallel execution has setup overhead (~100–500 ms). Rule of thumb:

- **Worth it**: ≥ 10 symbols *and* ≥ 1 000 bars each.
- **Not worth it**: small datasets (< ~10 symbols or < ~1 000 bars each) —
  sequential is usually faster.

## 8. Extending the package — `register_metric()`

Register a custom metric that returns the mean absolute deviation:

```{r register}
register_metric(
  name    = "mad_metric",
  fn      = function(x, .args = list()) {
    c(mad_metric = mean(abs(x - mean(x, na.rm = TRUE)), na.rm = TRUE))
  },
  outputs = "mad_metric"
)

available_metrics()

mad_out <- roll_nonlinear(
  dt[symbol == "AAPL"],
  metrics     = "mad_metric",
  window_size = 64L,
  stride      = 16L
)
head(mad_out[, .(time, mad_metric)])
```

Registered metrics work identically under both sequential and parallel
execution.

## Caveats and known limitations

- **Lyapunov & correlation dimension** require substantial data (≥ 500–1 000
  bars per window) to produce meaningful estimates. Short windows yield
  unreliable values; use them with `\donttest{}` guards in your scripts.
- **Sample/approximate entropy** with `r_frac` is volatility-normalised by
  default. Pass `r` explicitly if you need volatility-sensitive comparisons.
- **Parallel overhead**: for small datasets, sequential mode is faster.
  Prefer `parallel = TRUE` for wide, large-bar symbol sets.
